Methods for Large Scale Learning to Rank

### 2014-11-13 report/Chapters/RelatedWork.tex
meer Dolf

Acharyya et al. \cite{Acharyya2012, Acharyya2013} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{NDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}. No experiments on speed-up have been performed.
\subsection{Parallel robust Learning to Rank}
Robust learning \cite{Huber1981} is defined as the task to lean a model in the presence of outliers. Yun et al. described a \cite{Yun2014} robust Learning to Rank model called RoBiRank that has the additional advantage that it is executable in parallel. RoBiRank uses parameter expansion to linearise a surrogate loss function, after which the elements of the linear combination can be divided over the available nodes. The Parameter expansion trick was proposed by Gopal and Yang \cite{Gopal2013}, who originally proposed it for multinomial logistic regression. Unfortunately, no speed-up experiments were mentioned for the RoBiRank method, since Yun et al. focussed their research more on robust ranking than on parallel ranking. The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel Rank\ac{SVM} \cite{Herbrich1999, Joachims2002} described in section \ref{sec:related_work_low_complexity}.
\subsection{Distributed Stochastic Gradient Descent}
Long et al. \cite{Long2012} described a special case of their pairwise cross-domain factor Learning to Rank method using distributed optimization of \acs{SGD}-based on Hadoop MapReduce. Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce-based method described by  Zinkevich et al. \cite{Zinkevich2010} has been used. Real world data from Yahoo! has been used to show that the model is effective. Unfortunately the speed-up obtained by training their model in parallel is not reported.\\

\section{Ensemble learning for parallel Learning to Rank}
Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}. The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods. Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (often called stacking) \cite{Wolpert1992}.\\

@@

Ye et al. \cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both \ac{MPI} and Hadoop. Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting. Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}. Experiments showed the Hadoop implementation to resulted in too expensive communication cost to be useful. Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm. The \ac{MPI} approach proved to be successful and obtained near linear speed-ups.\\

Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner. Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the trees in the LambdaMART tree-ensemble are split instead of the data. The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data. Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.

@@

Yandex researchers Pavlov et al. \cite{Pavlov2010} were the first to proposed a boosting-wrapped-in-bagging approach, which they called BagBoo. The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\

@@

\subsection{Stacked generalisation}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}. \ac{DSN}s are based on the stacked generalization (also known as stacking) ensemble learning method \cite{Wolpert1992}, which is a multi-layer approach where multiple learning models are stacked on top of each other in such a way that the outputs of the models are used as the input features for the models in the next layer of the stack. Stacked generalisation models are easy to compute in parallel, as the models within the same layer can be trained independently. Deng et al. \cite{Deng2013} used a deep architecture of stacked generalisation (\ac{DSN}) to make pointwise relevance label predictions such that the \ac{MSE} of the predicted relevance labels as compared to the ground truth relevance labels are optimised. No speed-up results of parallel computation are reported, although it is stated repeatedly that the method is scalable and easy to compute in parallel

>>>>

Acharyya et al. \cite{Acharyya2012, Acharyya2013} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{NDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}. No experiments on speed-up have been performed.
\subsection{Parallel robust Learning to Rank}
Robust learning \cite{Huber1981} is defined as the task to lean a model in the presence of outliers. Yun et al. described a \cite{Yun2014} robust Learning to Rank model called RoBiRank that has the additional advantage that it is executable in parallel. RoBiRank uses parameter expansion to linearise a surrogate loss function, after which the elements of the linear combination can be divided over the available nodes. The Parameter expansion trick was proposed by Gopal and Yang \cite{Gopal2013}, who originally proposed it for multinomial logistic regression. Unfortunately, no speed-up experiments were mentioned for the RoBiRank method, since Yun et al. focussed their research more on robust ranking than on parallel ranking. The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel Rank\ac{SVM} \cite{Herbrich1999, Joachims2002} described in section \ref{sec:related_work_low_complexity}.
\subsection{Distributed Stochastic Gradient Descent}
Long et al. \cite{Long2012} described a special case of their pairwise cross-domain factor Learning to Rank method using distributed optimization of \acs{SGD}-based on Hadoop MapReduce. Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce-based method described by  Zinkevich et al. \cite{Zinkevich2010}. Real world data from Yahoo! has been used to show that the model is effective. Unfortunately the speed-up obtained by training their model in parallel is not reported.\\

\section{Ensemble learning for parallel Learning to Rank}
Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}. The invention of the boosting method resulted in an increasing focus on ensemble methods, which are methods of combining multiple weak models. Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (often called stacking) \cite{Wolpert1992}.\\

@@

Ye et al. \cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both \ac{MPI} and Hadoop. Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting. Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}. Experiments showed the Hadoop implementation to result in too high communication cost to be useful. Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm. The \ac{MPI} approach proved to be successful and obtained near linear speed-ups.\\

Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the \mbox{LambdaMART} \cite{Wu2008} model in a distributed manner. Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the trees in the LambdaMART tree-ensemble are split instead of the data. The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data. Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.

@@

Yandex researchers Pavlov et al. \cite{Pavlov2010} were the first to propose a boosting-wrapped-in-bagging approach, which they called BagBoo. The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\

@@

\subsection{Stacked generalisation}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}. \ac{DSN}s are based on the stacked generalization (also known as stacking) ensemble learning method \cite{Wolpert1992}, which is a multi-layer approach where multiple learning models are stacked on top of each other in such a way that the outputs of the models are used as the input features for the models in the next layer of the stack. Stacked generalisation models are easy to compute in parallel, as the models within the same layer can be trained independently. Deng et al. \cite{Deng2013} used a deep architecture of stacked generalisation (\ac{DSN}) to make pointwise predictions of relevance labels such that the \ac{MSE} of those predictions compared to the ground truth relevance labels are minimised. No speed-up results of parallel computation are reported, although it is stated repeatedly that the method is scalable and easy to compute in parallel.
